---
title: "Palmer Penguins"
author: "Antonia M."
date: "22 4 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

## Palmer Penguins dataset

This R Markdown document is connected to analysis of the Palmer Penguins dataset.
In following pages, we will perform Principal Component Analysis in order to explore 
various parameters, e.g. bill length and depth. Firstly, the dataset was loaded 
to R Markdown and the structure of the datasets was examined.


```{r dataset loading, echo = FALSE, warning=FALSE}
library(ggplot2)
library(palmerpenguins)
data(package = 'palmerpenguins')
str(penguins)

```

## Visualization of the penguin mass and flipper length correlation

```{r mass vs flip len}
mass_flipper <- ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for Adelie, Chinstrap and Gentoo Penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.2, 0.7),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

mass_flipper
```

## Visualization of the flipper length vs bill length correlation 

```{r flip vs bill length}
flipper_bill <- ggplot(data = penguins,
                         aes(x = flipper_length_mm,
                             y = bill_length_mm)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper and bill length",
       subtitle = "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)",
       y = "Bill length (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.85, 0.15),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

flipper_bill
```




## Visualization of the bill length and depth correlation

You can also embed plots, for example:

```{r plot bill_len_dep, echo=FALSE, warning=FALSE}
bill_len_dep <- ggplot(data = penguins,
                         aes(x = bill_length_mm,
                             y = bill_depth_mm,
                             group = species)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, aes(color = species)) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin bill dimensions",
       subtitle = "Bill length and depth for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Bill length (mm)",
       y = "Bill depth (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.85, 0.15),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

bill_len_dep
```

Next, we perform the analysis of the Simpson's paradox. 

## Simpson's Paradox visualization

```{r simpson, warning=FALSE}
bill_no_species <- ggplot(data = penguins,
                         aes(x = bill_length_mm,
                             y = bill_depth_mm)) +
  geom_point() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin bill dimensions (omit species)",
       subtitle = "Palmer Station LTER",
       x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme(plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot") +
  geom_smooth(method = "lm", se = FALSE, color = "gray50")

bill_no_species
```

To conclude, here is the information about the session:

```{r session info}
sessionInfo()
```

